---
title: "`r params$new_title` tree cover change and rangeland production report"
author: 'Report derived from the Rangeland Analysis Platform'
params:
  new_title: params$new_title
  main_df: params$main_df
  cty_sf: params$cty_sf
  sta_sf: params$sta_sf
  sta_ctys_sf: params$sta_ctys_sf
  current_year: params$current_year
  rap_url: params$rap_url
  type: params$type
output: pdf_document
always_allow_html: true

---

```{r setup, include=FALSE}
library(here)
library(tidyverse)
library(sf)
library(knitr)
require(scales)
library(reshape2)
library(leaflet)
library(leaflet.extras)
library(kableExtra)
library(patchwork)

knitr::opts_chunk$set(echo = TRUE, root.dir = here())
```

<div style="margin-bottom:30px;">
</div>
  
## `r paste(params$type, ' map')` 
Click any county on the map below to access its report.  
```{r echo = FALSE, warning = FALSE, fig.width=10}
scale_range <- c(-1, 1)

pal <- colorNumeric("RdBu", domain = scale_range)

zoom_level = params$sta_sf$zoom %>% unlist()

if (params$type == 'County'){
  leaflet(options = leafletOptions(zoomControl = FALSE,  minZoom = zoom_level, maxZoom = zoom_level, dragging = FALSE, attributionControl=FALSE)) %>%
    addPolygons(data = params$sta_sf, 
                fillColor = "#00000000", 
                color = '#5e5e5e', 
                opacity = 1, 
                weight = 4) %>%
     addPolygons(data = params$sta_ctys_sf, 
                 fillColor = ~pal(is10uw), 
                 color = '#5e5e5e', 
                 weight = 1.5, 
                 opacity = 1, 
                 label = ~NAMELSAD, 
                 popup = paste(sta_ctys_sf$NAMELSAD, "<BR>", "<a href =   '",sta_ctys_sf$up_path, "' target = '_self'>", 'See the report </a>'),
                 labelOptions = labelOptions(style = list(
                                               "box-shadow" = "3px 3px rgba(0,0,0,0.25)",
                                               "font-size" = "16px")))  %>%
    addPolygons(data = params$cty_sf, 
                fillColor = "#00000000", 
                color = '#FF0000', 
                weight = 4, 
                opacity = 1, 
                fillOpacity = 1, 
                stroke = TRUE, 
                smoothFactor = 0.5,) %>%
     addLegend(values = scale_range, 
               group = "circles", 
               title = 'Rangeland area trend due </br>to tree cover change', 
               position = "bottomleft", 
               colors = c('#849bb4', '#a2c4dd', '#d9eaf3', '#ffffff','#fcded1', '#de9d9f', '#b88291'),
               labels= c("Increase", "", "", "", "", "", "Decrease")) %>%
    setMapWidgetStyle(list(background= "white")) 
  
} else if (params$type == 'State') {
    leaflet(options = leafletOptions(zoomControl = FALSE,  minZoom = zoom_level, maxZoom = zoom_level, dragging = FALSE, attributionControl=FALSE)) %>%
     addPolygons(data = params$sta_sf, 
                 fillColor = "#00000000", 
                 color = '#5e5e5e', 
                 opacity = 1, 
                 weight = 4) %>%
     addPolygons(data = params$sta_ctys_sf, 
                 fillColor = ~pal(is10uw), 
                 color = '#5e5e5e', 
                 weight = 1.5, 
                 opacity = 1, 
                 label = ~NAMELSAD, 
                 popup = paste(sta_ctys_sf$NAMELSAD, "<BR>", "<a href =   '",sta_ctys_sf$dn_path, "' target = '_self'>", 'See the report </a>'),
                 labelOptions = labelOptions(style = list(
                                               "box-shadow" = "3px 3px rgba(0,0,0,0.25)",
                                               "font-size" = "16px")))  %>%
     addLegend(values = scale_range, 
               group = "circles", 
               title = 'Rangeland area trend due </br>to tree cover change', 
               position = "bottomleft", 
               colors = c('#849bb4', '#a2c4dd', '#d9eaf3', '#ffffff','#fcded1', '#de9d9f', '#b88291'),
               labels= c("Increase", "", "", "", "", "", "Decrease")) %>%
     setMapWidgetStyle(list(background= "white"))
  }
```


|
|
## Summary statistics for `r params$new_title`  
#### State and county summary  
_  

```{r echo = FALSE, warning = FALSE}
char_df <- tibble(init = 10)
char_df[, 'Total area'] = paste(as.character(round(params$main_df$totalArea[1]), 0), 'acres')
char_df[, 'Analysis area'] <- paste(as.character(round(params$main_df$analysisArea[1]), 0), 'acres')
char_df[, 'Area excluded from analysis'] <- paste(as.character(round(params$main_df$totalArea[1] - params$main_df$analysisArea[1]), 0), 'acres')
char_df <- char_df[,2:4]

char_df <- char_df %>%
  t() %>%
  melt() %>%
  mutate(Value = prettyNum(value, big.mark = ","),
         Blank = c('','','')) %>%
  select(c(Variable = Var1, Value,Blank))

kable(char_df, col.names = NULL, booktabs = FALSE) %>%
  column_spec(1, width = "15cm") %>%
  column_spec(2, width = "5cm") %>%
  column_spec(3, width = "5cm") %>%
  row_spec(1:3, hline_after = TRUE) %>%
  kable_styling()
```



#### Tree cover and herbaceous production summary  
_  

```{r echo = FALSE, warning = FALSE}
# Statistics for report
figure_v <- c("[Figure 1](#tree-cover-and-herbaceous-production)",
              "[Figure 2](#tree-cover-and-herbaceous-production)", 
              "", 
              "", 
              "[Figure 3](#herbaceous-production-gainslosses-resulting-from-tree-cover-change)", 
              "[Figure 4](#herbaceous-production-gainslosses-resulting-from-tree-cover-change)")
main_df_rec <- params$main_df %>%
  filter(year == params$current_year) 

stat_df <- tibble(init = 10)
stat_df[, paste('Tree cover in', as.character(params$current_year))] <- paste(as.character(round(main_df_rec$treeArea,0)), 'acres')
stat_df[, paste('Tree cover change since 1990')] <- paste(as.character(round((tail(params$main_df$treeArea, n = 1)) - (head(params$main_df$treeArea, n = 1)), 0)), 'acres')
stat_df[, paste('Tree cover percent in', as.character(params$current_year))] <- paste(as.character(round(main_df_rec$treeCover,2)),'%', sep = '')
stat_df[, paste('Herbaceous production in', as.character(params$current_year))] = paste(as.character(round(main_df_rec$biomass,0)), 'tons')
stat_df[, paste('Herbaceous production gains/losses resulting from tree cover change in', as.character(params$current_year))] <- paste(as.character(-round(main_df_rec$yieldgap,0)), 'tons')
stat_df[, paste('Herbaceous production gains/losses resulting from tree cover change since 1990')] <- paste(as.character(-round(sum(params$main_df$yieldgap),0)), 'tons') 
stat_df <- stat_df[,2:7]

stat_df <- stat_df %>%
  t() %>%
  melt() %>%
  mutate(Figure = figure_v[row_number()],
         Value = prettyNum(value, big.mark = ",")) %>%
  select(c(Variable = Var1, Value, Figure))

kable(stat_df, col.names = NULL) %>%
  column_spec(1, width = "15cm") %>%
  column_spec(2, width = "5cm") %>%
  column_spec(3, width = "5cm") %>%
  row_spec(1:6, hline_after = TRUE) %>%
  kable_styling()
```

<div style="margin-bottom:15px;">
</div>

|
|
## Tree cover and herbaceous production plots for `r params$new_title`  
#### Tree cover and herbaceous production  
_  

```{r echo = FALSE, fig.width=10}
# Annual production plot
prod_plot <- ggplot()+
  geom_line(params$main_df, mapping = aes(x = year, y = biomass), color = 'grey20', size = 1)+
  geom_point(params$main_df, mapping = aes(x = year, y = biomass), color = 'grey20', size = 2)+
  labs(title = 'Figure 1: Herbaceous production', y = 'Annual production (tons)')+
  scale_y_continuous(labels = comma, limits = c(0, (max(params$main_df$biomass))*1.1))+
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  theme_minimal()+
  theme(axis.title.x=element_blank(),
        plot.title.position = "plot")

# Tree cover area
tree_plot <- ggplot()+
  geom_line(params$main_df, mapping = aes(x = year, y = treeArea), color = '#0E8E00', size = 1)+
  geom_point(params$main_df, mapping = aes(x = year, y = treeArea), color = '#0E8E00', size = 2)+
  geom_hline(yintercept = head(params$main_df$treeArea, n=1), linetype = 'dashed', color = 'grey30', size = 1)+
  labs(title = 'Figure 2: Tree cover', y = 'Tree cover (acres)')+
  scale_y_continuous(labels = comma, limits = c(0, max(params$main_df$treeArea)*1.1))+
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  theme_minimal()+
  #lims(y = c(0,max(params$main_df$treeArea)*1.1))+
  theme(axis.title.x=element_blank(),
        plot.title.position = "plot")

(prod_plot | tree_plot)
```

<div style="margin-bottom:50px;">
</div>

#### Herbaceous production gains/losses resulting from tree cover change  
_  

```{r echo = FALSE, fig.width=10}
annYG_plot <- ggplot()+
  geom_line(params$main_df, mapping = aes(x = year, y = -yieldgap), color = 'red', size = 1)+
  geom_point(params$main_df, mapping = aes(x = year, y = -yieldgap), color = 'red', size = 2)+
  geom_hline(yintercept = 0, linetype = 'dashed', color = 'grey30', size = 1)+
  labs(title = 'Figure 3: Annual herbaceous production gains/losses\nresulting from tree cover change', y = 'Production gains/losses (tons)')+
  coord_cartesian(ylim = c(-(max(abs(params$main_df$yieldgap)))*1.1,(max(abs(params$main_df$yieldgap)))*1.1))+
  scale_y_continuous(labels = comma)+
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  theme_minimal()+
  theme(axis.title.x=element_blank(),
        plot.title.position = "plot")

cumYG_plot <- ggplot()+
  geom_line(params$main_df, mapping = aes(x = year, y = -cumYieldGap), color = 'red', size = 1)+
  geom_point(params$main_df, mapping = aes(x = year, y = -cumYieldGap), color = 'red', size = 2)+
  geom_hline(yintercept = 0, linetype = 'dashed', color = 'grey30', size = 1)+
  labs(title = 'Figure 4: Cumulative herbaceous production gains/losses\nresulting from tree cover change', y = 'Production gains/losses (tons)')+
  coord_cartesian(ylim = c((-max(abs(params$main_df$cumYieldGap)))*1.1,(max(abs(params$main_df$cumYieldGap)))*1.1))+
  scale_y_continuous(labels = comma)+
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  theme_minimal()+
  theme(axis.title.x=element_blank(),
        plot.title.position = "plot")

(annYG_plot | cumYG_plot)
```

<div style="margin-bottom:50px;">
</div>
  
```{r echo = FALSE, warning = FALSE, fig.width=10}
plot_df <- params$main_df %>%
  select(c(year, `At-risk` = classAtRisk, Intact = classIntact)) %>%
  melt(id = 'year')

ggplot()+
  geom_bar(plot_df, mapping = aes(fill = variable, x = year, y = value), position = position_dodge(width = 0.9), stat = "identity")+
  labs(y = 'Cover (acres)', title = 'Figure 5a: Intact and at-risk area')+
  scale_fill_manual(
        values = c(
          `At-risk` = "#FDE64B",
          Intact = "#4F963C"))+
  scale_y_continuous(labels = comma)+
  theme_minimal()+
  theme(legend.title = element_blank(),
        legend.position="bottom",
        axis.title.x=element_blank(),
        plot.title.position = "plot")
```

<div style="margin-bottom:35px;">
</div>

## Reference
Access the complete data that from which this report is derived: 

* [View tree cover data in the Rangeland Analysis Platform web application](`r params$rap_url`){target="_blank"}  
* [Download complete tree cover change time-series CSV](`r ifelse(params$type == 'County', './County.csv', './State.csv')`)  

  
*Report generated on `r Sys.Date()`.*

<div style="margin-bottom:35px;">
</div>